Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2017-10-17 TACAS2018/paper.tex
sec 6.2

While in the previous section, we assumed that
${\mat{A}_d}_{\langle C \rangle}=\mat{A}_d$, ${\mat{B}_d}_{\langle C \rangle}=\mat{B}_d$ and ${\mat{C}_d}_{\langle C \rangle}=\mat{C}_d$ for a given time discretization, in this section
we take into consideration the numerical representation and namely the fact that the domain for both
the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.

%% Thus far we looked at checking stability
%% %have presented a verification procedure for
%% of a continuous plant in discretized time domain (i.e., the range of the quatizers is still
%% $\mathbb{R}$).  The next step we must realize is that the controller and the
%% observer exist in a digital program, hence their domain is
%% $\mathbb{R}_{\langle C \rangle}$.

This has two separate effects on the
model.  If we recall Equation~\eqref{eq:observer_LTI}, the matrices of the observer
and the controller are not necessarily those of the plant.  In fact we have
$\mat{A},\mat{B},\mat{C} \in \mathbb{R}^*$, ${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},{\mat{C}_d}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$ where $*$ represents the appropriate
dimension for each matrix.  Unfortunately this means that although combined
matrices such as $\mat{A}_t$ are in the reals, they suffer from the fixed
word restrictions in their composing matrices.

When our observer is implemented using a finite precision controller, new
problems arise due to the FWL effects.  Let ${\mat{A}_d}_{\langle C
\rangle}$, ${\mat{B}_d}_{\langle C \rangle}$ and ${\mat{C}_d}_{\langle C
\rangle}$ be the dynamics represented in $\mathbb{R}_{\langle C\rangle}^*$,
%
%and $\mat{A}_e=\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}$, $\mat{B}_e=\mat{B}_d-{\mat{B}_d}_{\langle C \rangle}$,
%and $\mat{C}_e=\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}$ their corresponding errors.  
Equation \eqref{eq:observer_LTI_cf} becomes:
%
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}_{c} \vec{r}_k
+\mat{B}_{\eta}\left [\begin{array}{c}\vec{\eta}_2\\ \vec{\eta}_1\end{array}\right]\nonumber\\
\vec{y}_k=&\mat{C}_{t}\vec{x}_k'% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
where
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}\right)\mat{T}_c^{-1}&
\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
\mat{C}_{e}=&\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}\\
\mat{T}_c=&\mat{R}_{cf}\mat{R}^{-1}&
\mat{T}_o=&\mat{W}_o^{-1}\mat{W}_{of}\\
\mat{C}_t=&\mat{C}_d\mat{T}_c&
\mat{L}_{\langle C \rangle}=& \mat{T}_o^{-1}\left[ \begin{array}{ccc}l_1&\cdots&l_n\end{array}\right]^T
\end{align*}
}
{\setlength{\abovedisplayskip}{0pt}
\begin{align*}
\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
\mat{B}_{c}=&\left[\begin{array}{c}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{k}_r\\\mat{0}\end{array}\right]&
\mat{B}_{\eta}=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]
\end{align*}
}
The first effect of this new equation is that $\mat{A}'$ is no longer upper
diagonal.  We must therefore compensate for the errors introduced in order
to find a stable solution.  We can use Gershgoring disks to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(|{\mat{A}_e}_{i,*}|_2)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, n],
|{\lambda_o}_{i}-\hat{\lambda_o}_{i}|<\delta_o$,where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer and $\hat{\lambda_o}_{i}$ the eigenvalues of
the perturbed observer.  We note that in the case of stability we only care
about the error with respect to the unit circle, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence given a
desired upper value $\overline{\lambda}$, we apply Jury's criterion on the
polynomial.
%
\begin{equation}
P_o'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j : \overline{\lambda'}=(1+\delta_o)\overline{\lambda}
\end{equation}
%
In the case of the reachable eigenvalues, the procedure is the same except
that we obtain our bound $\delta_c=\max(|{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}|_2)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_r'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j : \overline{\lambda'}=(1+\delta_c)\overline{\lambda}
\end{equation}

As a final annotation, we remark that the safety specification must be
applied to the model described in this section, accounting for the
representation errors of the FWL domain.

@@

We are not limited to the synthesis of digital stabilizing controllers -- a
well known task in the literature on digital control systems -- but target
safety requirements with an overall approach that is sound and automated. 
More specifically, we require that the closed-loop system meets given safety
specifications.  A safety specification gives raise to a requirement on the
states of the model, such that the feedback controller (namely the choice of
the gains matrix~$K$) must ensure that the state never violates the
requirement.  Note that a stable, closed-loop system is not necessarily a
safe system: indeed, the state values may leave the safe part of the state
space while they converge to the equilibrium, which is typical in the case
of oscillatory dynamics.  In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}\equiv& \underline{y} \leq y_k \leq \overline{y}
\iff \forall k\geq 0.\, \mat{G}\vec{x}_k \leq \vec{h},
\label{eq:safetyspec}
\end{align}

where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe the safe state-space.

@@

%==========

%% Now that we have established convergence, we may examine the safety
%% specification of the state space.
In the case of SISO systems, the safety specification is
given by a range of the input and a range of the output which are considered
safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in
[\underline{y}\ \ \overline{y}]$)
Whilst there are other properties we may
include in the verification, we will solely focus on these for simplicity of
the explanation.  We assume that we do not want saturation in our
system, hence the control loop has only one mode.
%We have also identified the quantization noise as a bounded non-deterministic input.
First we must
translate the safety specification into the statespace.  Replacing on the
second part of equation \eqref{eq:observer_LTI_cf} we obtain
%
\begin{equation}
\phi_{safety} := \underline{y}<\mat{C}\vec{x}_k<\overline{y}, \forall k \geq 0
\label{eq:of_bounds}
\end{equation}
\todo{Using \eqref{eq:observer_LTI_cf} doesn't result in this.}
%
\todo{transient response?}
%
%Note that we use the bounds of the input after the transformation by $\mat{B}$ since the matrix may reverse the bounds. 
%
%We evaluate the progression of the system over time by calculating
%$\vec{x}_k$ to validate against this specification.  For this step we require
%an initial set $\vec{x}_0$ from which the system progresses.  Obviously, this set
%must satisfy the specification, otherwise the system will be
%unsafe to begin with.  For efficiency purposes, the inequality in \eqref{eq:of_bounds} is transformed into reachable form using $\mat{T}_c, \mat{T}_o$ which reduces the complexity from $\mathcal{O}(p^2)$ to $\mathcal{O}(p)$.

>>>>

In this section, we assume that the observer model is further discretised in
the value domain, i.e., the domain for both
the controller and the observer is $\mathbb{R}_{\langle C \rangle}$.
%% While in the previous section, we assumed that
%% ${\mat{A}_d}_{\langle C \rangle}=\mat{A}_d$, ${\mat{B}_d}_{\langle C \rangle}=\mat{B}_d$ and ${\mat{C}_d}_{\langle C \rangle}=\mat{C}_d$ for a given time discretization, in this section
%% we take into consideration the numerical representation and namely
%% Thus far we looked at checking stability
%% %have presented a verification procedure for
%% of a continuous plant in discretized time domain (i.e., the range of the quatizers is still
%% $\mathbb{R}$).  The next step we must realize is that the controller and the
%% observer exist in a digital program, hence their domain is
%% $\mathbb{R}_{\langle C \rangle}$.
%
%% This has two separate effects on the
%% model.  If we recall Equation~\eqref{eq:observer_LTI}, the matrices of the observer
%% and the controller are not necessarily those of the plant.  In fact we have
%% $\mat{A},\mat{B},\mat{C} \in \mathbb{R}^*$,
This means that we have 
${\mat{A}_d}_{\langle C
\rangle},{\mat{B}_d}_{\langle C \rangle},{\mat{C}_d}_{\langle C \rangle},
\mat{K}_{\langle C \rangle}, \mat{L}_{\langle C \rangle} \in
\mathbb{R}_{\langle C \rangle}^*$ where $*$ represents the appropriate
dimension for each matrix.  Unfortunately this means that although combined
matrices such as $\mat{A}_t$ are in the reals, they suffer from the fixed
word restrictions in their composing matrices.

When our observer is implemented using a finite precision controller, new
problems arise due to the FWL effects.  Let ${\mat{A}_d}_{\langle C
\rangle}$, ${\mat{B}_d}_{\langle C \rangle}$ and ${\mat{C}_d}_{\langle C
\rangle}$ be the dynamics represented in $\mathbb{R}_{\langle C\rangle}^*$,
%
%and $\mat{A}_e=\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}$, $\mat{B}_e=\mat{B}_d-{\mat{B}_d}_{\langle C \rangle}$,
%and $\mat{C}_e=\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}$ their corresponding errors.  
Equation \eqref{eq:observer_LTI_cf} becomes:
%
\begin{align}
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_{k+1}
=& \left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]
\left [\begin{array}{c}\vec{x}'\\ \vec{x}_e' \end{array}\right]_k
+\mat{B}_{c} \vec{r}_k
+\mat{B}_{\eta}\left [\begin{array}{c}\vec{\eta}_2\\ \vec{\eta}_1\end{array}\right]\nonumber\\
\vec{y}_k=&\mat{C}_{t}\vec{x}_k'% + \mat{D}_{d}\vec{u}_k
\label{eq:observer_LTI_FWL_cf}
\end{align}
where
{\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{align*}
\mat{A}_{c}=&\mat{T}_c\left(\mat{A}_d-\mat{B}_d\mat{K}\right)\mat{T}_c^{-1}&
\mat{A}_{o}=&\mat{T}_o\left({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle}\right)\mat{T}_o^{-1}\\
\mat{B}_{o}=&\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}&
\mat{B}_{e}=&\mat{T}_c(\mat{B}_d-{\mat{B}_d}_{\langle C \rangle})\mat{K}_{\langle C \rangle}\mat{T}_o^{-1}\\
\mat{A}_{e}=&\mat{T}_o\left(\mat{A}_d-{\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}\mat{C}_e\right)\mat{T}_o^{-1}&
\mat{C}_{e}=&\mat{C}_d-{\mat{C}_d}_{\langle C \rangle}\\
\mat{T}_c=&\mat{R}_{cf}\mat{R}^{-1}&
\mat{T}_o=&\mat{W}_o^{-1}\mat{W}_{of}\\
\mat{C}_t=&\mat{C}_d\mat{T}_c&
\mat{L}_{\langle C \rangle}=& \mat{T}_o^{-1}\left[ \begin{array}{ccc}l_1&\cdots&l_n\end{array}\right]^T
\end{align*}
}
{\setlength{\abovedisplayskip}{0pt}
\begin{align*}
\mat{A}'=&\left [\begin{array}{cc}\mat{A}_{c}+\mat{B}_{e}&\mat{B}_{o}\\ \mat{A}_{e}&\mat{A}_{o}\end{array}\right]&
\mat{B}_{c}=&\left[\begin{array}{c}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}\mat{k}_r\\\mat{0}\end{array}\right]&
\mat{B}_{\eta}=&\left[\begin{array}{cc}\mat{T}_c{\mat{B}_d}_{\langle C \rangle}&\mat{0}\\\mat{0}&-\mat{T}_o\mat{L}_{\langle C \rangle}\end{array}\right]
\end{align*}
}
The first effect of this new equation is that $\mat{A}'$ is no longer upper
diagonal.  We must therefore compensate for the errors introduced in order
to find a stable solution.  We can use Gershgoring \todo{citation} disks to limit the error
in the expected eigenvalues.  Let $\delta_o=\max(|{\mat{A}_e}_{i,*}|_2)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, n],
|{\lambda_o}_{i}-\hat{\lambda_o}_{i}|<\delta_o$,where ${\lambda_o}_{i}$ are the
true eigenvalues of the observer and $\hat{\lambda_o}_{i}$ the eigenvalues of
the perturbed observer.  We note that in the case of stability we only care
about the error with respect to the unit circle, which is equivalent to
upper bounding the estimation of the largest eigenvalue.  Hence, given a
desired upper value $\overline{\lambda}$, we apply Jury's criterion on the
polynomial.
%
\begin{equation}
P_o'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j : \overline{\lambda'}=(1+\delta_o)\overline{\lambda}
\end{equation}
%
In the case of the reachable eigenvalues, the procedure is the same except
that we obtain our bound $\delta_c=\max(|{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}|_2)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_r'=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}s^j : \overline{\lambda'}=(1+\delta_c)\overline{\lambda}
\end{equation}

%% As a final annotation, we remark that the safety specification must be
%% applied to the model described in this section, accounting for the
%% representation errors of the FWL domain.

@@

%% Now that we have established convergence, we may examine the safety
%% specification of the state space.
A safety specification gives raise to a requirement on the
states of the model, such that the controller and the observer (namely the 
matrices~$K$ and~$L$) must ensure that the state never violates the
requirement.

In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}\equiv& \underline{y} \leq y_k \leq \overline{y}
\iff \forall k\geq 0.\, \mat{G}\vec{x}_k \leq \vec{h},
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe the safe state-space.
%
%% \begin{align}
%% \phi_\mathit{safety}\equiv& \underline{y} \leq y_k \leq \overline{y}
%% \iff \forall k\geq 0.\, \mat{G}\vec{x}_k \leq \vec{h},
%% \label{eq:safetyspec}
%% \end{align}
%
%% For an observed system, the safety specification is
%% given by a constraint on the output which are considered
%% safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in
%% [\underline{y}\ \ \overline{y}]$)
%% Whilst there are other properties we may
%% include in the verification, we will solely focus on these for simplicity of
%% the explanation.  We assume that we do not want saturation in our
%% system, hence the control loop has only one mode.
%% First we must
%% translate the safety specification into the statespace.  Replacing on the
%% second part of equation \eqref{eq:observer_LTI_cf} we obtain
%% %
%% \begin{equation}
%% \phi_{safety} := \underline{y}<\mat{C}\vec{x}_k<\overline{y}, \forall k \geq 0
%% \label{eq:of_bounds}
%% \end{equation}
%
%% We are not limited to the synthesis of digital stabilizing controllers -- a
%% well known task in the literature on digital control systems -- but target
%% safety requirements with an overall approach that is sound and automated. 
%% More specifically, we require that the closed-loop system meets given safety
%% specifications.  A safety specification gives raise to a requirement on the
%% states of the model, such that the feedback controller (namely the choice of
%% the gains matrix~$K$) must ensure that the state never violates the
%% requirement.  Note that a stable, closed-loop system is not necessarily a
%% safe system: indeed, the state values may leave the safe part of the state
%% space while they converge to the equilibrium, which is typical in the case
%% of oscillatory dynamics.  In~this work, the safety property is expressed as:
%% %
%% \begin{align}
%% \phi_\mathit{safety}\equiv& \underline{y} \leq y_k \leq \overline{y}
%% \iff \forall k\geq 0.\, \mat{G}\vec{x}_k \leq \vec{h},
%% \label{eq:safetyspec}
%% \end{align}

%% where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe the safe state-space.

@@

%
%Note that we use the bounds of the input after the transformation by $\mat{B}$ since the matrix may reverse the bounds. 
%
%We evaluate the progression of the system over time by calculating
%$\vec{x}_k$ to validate against this specification.  For this step we require
%an initial set $\vec{x}_0$ from which the system progresses.  Obviously, this set
%must satisfy the specification, otherwise the system will be
%unsafe to begin with.  For efficiency purposes, the inequality in \eqref{eq:of_bounds} is transformed into reachable form using $\mat{T}_c, \mat{T}_o$ which reduces the complexity from $\mathcal{O}(p^2)$ to $\mathcal{O}(p)$.
